---
title: "R Notebook"
output: md_document
---

## August 10th: Functions for separating nodes into border and non border groups

```{r}
library(igraph)
library(ggplot2)
library(dplyr)
```

# Load data, calculate communities, set parameters for plotting 
Read in gene counts & weighted iGraph networks
```{r}
cell.type.distributions <- readRDS('../expression.vals.rds')
g04 <- readRDS('../g04.rds')
g08 <- readRDS('../g08.rds')
```

Save community objects using infomap and louvain:
```{r}
g04.infomap <- cluster_infomap(g04)
g04.louvain <- cluster_louvain(g04)

g08.infomap <- cluster_infomap(g08)
g08.louvain <- cluster_louvain(g08)
```

Create functions for consistent color scheme of cell types across different graphs and clustering results:
```{r}
#make list of any cell types over 15% for a given spot, and return the top (#1) for each community
comm.top.celltypes <- function(communities.object, cell.types=cell.type.distributions){
  x <- lapply(communities(communities.object),function(z){
    x <- colMeans(cell.types[z,])
    x <- x[sort.list(x,decreasing = TRUE)]
    if (length(which(x>.3))>0){
      x[which(x>.3)]
    } else if (length(which(x>.15))>0){
      x[which(x>.15)]
    } else {
      x[1:2]
    }
  })
  as.factor(sapply(x,function(i){names(i)[1]}))
}

#colors used for maximum number of all cell types
define.global.palette <- function(communities.object, cell.types=cell.type.distributions){
  global.pal <- topo.colors(length(unique(colnames(cell.types))),alpha=.2) #unique colorset for all available cell types
  names(global.pal) <- colnames(cell.types)
  global.pal
}

#colors used for this specific clustering (communities) object
define.local.pal <- function(communities.object, cell.types=cell.type.distributions){
  communities.object$global.palette[as.character(unique(communities.object$top.celltypes))] #unique colorset used for this instance 
}

#assign correct plot color to each community
define.community.colors <- function(communities.object, cell.types=cell.type.distributions){
  communities.object$local.palette[as.character(communities.object$top.celltypes)] 
}

set.community.attributes <- function(communities.object, subtitle='', cell.types=cell.type.distributions){
  communities.object$top.celltypes <- comm.top.celltypes(communities.object)
  communities.object$global.palette <- define.global.palette(communities.object)
  communities.object$local.palette <- define.local.pal(communities.object)
  communities.object$community.colors <- define.community.colors(communities.object)
  communities.object$subtitle <- subtitle
  communities.object
}
```

```{r}
g04.infomap <- set.community.attributes(g04.infomap, 'Infomap')
g04.louvain <- set.community.attributes(g04.louvain, 'Louvain')

g08.infomap <- set.community.attributes(g08.infomap, 'Infomap')
g08.louvain <- set.community.attributes(g08.louvain, 'Louvain')
```


Plot clustering results: https://igraph.org/python/doc/igraph.drawing-module.html
```{r}
plot.communities <- function(basegraph, communities.object) {
  plot(basegraph, sub = communities.object$subtitle
    ,mark.groups = communities.object, mark.expand = 2, edge.width=0.2 
         ,edge.color = "lightgrey"
         ,mark.col = communities.object$community.colors
         ,mark.border="black")
  legend("bottomright",inset=.02,title="Dominant cell type"
         ,fill=communities.object$local.palette
         ,legend=names(communities.object$local.palette) 
         )
}
```

```{r, fig.width = 7, fig.height = 7}
plot.communities(g04, g04.infomap)
plot.communities(g08, g08.infomap)

plot.communities(g04, g04.louvain)
plot.communities(g08, g08.louvain)
```

# Save list of nodes denoted as border regions for later enrichment analysis
*How does `membership()` handle nodes that are in multiple communities though?* Membership only assigns the 'best fit' (one label per node).

```{r}
find_true_border_edges <- function(basegraph, communities.object){
  #returns nx2 list of nodes that cross communities w different top cell types
  border.edges <- crossing(communities.object, basegraph)
  border.crossings <- ends(basegraph, E(basegraph)[border.edges]) #647
  print (paste("total number of edges:",length(border.edges))) #4019 (total number of edges)
  print (paste("community crossings:",sum(border.edges))) #647 edges that connect different communities
  #dim(border.crossings) #647 by 2
  #border.crossings[1:5,] #
  
  memb <- communities.object$top.celltypes[membership(communities.object)]
  
  edge.connects.different.dominant.celltypes <- memb[border.crossings[,1]] != memb[border.crossings[,2]]
  #length(edge.connects.different.dominant.celltypes)
  #sum(edge.connects.different.dominant.celltypes) #561 are true borders
  true.crossings <- border.crossings[edge.connects.different.dominant.celltypes,]
  print(paste("true community crossings:",dim(true.crossings)[1]))
  true.crossings
}

label_true_border_edges <- function(communities.object, true_border_edges){
  #returns nx2 list of top cell types for each node in the true border edges list
  memb <- communities.object$top.celltypes[membership(communities.object)] #top cell type of each community
  y <- memb[true_border_edges]
  dim(y) <- dim(true_border_edges)
  y
}

summarize_true_border_edges <- function(y){#y is a factor matrix from label_true_border_edges
  #sort each row, so visualizations on the true border edges can be created
  celltype.border.edges <- as.character(y)
  dim(celltype.border.edges) <- dim(y)
  celltype.border.edges <- t(apply(celltype.border.edges,1,sort))
  celltype.border.edges <- apply( celltype.border.edges , 1 , paste , collapse = " --- " )
  
  result <- as.data.frame(celltype.border.edges) %>%
    group_by(celltype.border.edges) %>%
    summarise(edge.count = n()) %>%
    arrange(-1*edge.count)
  result
}
```

```{r}
g08.infomap.border_edges <- find_true_border_edges(g08, g08.infomap)
y2 <- label_true_border_edges(g08.infomap, g08.infomap.border_edges)
#dim(y2)
y2[1:15,]
```

It looks like Astro-L2/3IT is the most common cross-community border.
```{r}
summarize_true_border_edges(y2) #would be cool if I added columns with the # of A spots & number of B spots too (non border spots)
```

# Code for selecting a specific cell type edge and target cell type to study
```{r}
astro.rows <- y2[,1]=='Astro' | y2[,2]=='Astro'
sum(astro.rows) #167 total
l2l3it.rows <- y2[,1]=='L2/3 IT' | y2[,2]=='L2/3 IT'
sum(l2l3it.rows)

sum(astro.rows & l2l3it.rows) #146 astro-l2l3it edges

memb <- g08.infomap$top.celltypes[membership(g08.infomap)] #membership of every single spot
cases <- unique(as.numeric(g08.infomap.border_edges[astro.rows & l2l3it.rows,])) #unique nodes in the 146 astro-l2l3it edges
cases <- cases[memb[cases]=='Astro'] #only pull the astro nodes in astro-l2leit edges (59)
all.astro <- which(memb=='Astro')
```

Identify astro spots in astro-L2/L3 IT edges. And then identify controls (astro spots that are not in any other cross community edge, i.e. an interior spot).
```{r}
length(all.astro)
length(cases)
non.border.controls <- all.astro[!all.astro %in% as.numeric(g08.infomap.border_edges)] #for control (non border spots), remove any in our target
length(non.border.controls)
```

```{r}
case.control.colors <- rep(1,1072)
case.control.colors[cases] <- 2
case.control.colors[non.border.controls] <- 3
case.control.colors <- as.factor(-1*case.control.colors)

case.control.size <- rep(1,1072)
case.control.size[cases] <- 3
case.control.size[non.border.controls] <- 3
```

```{r}
plot.cases.and.controls <- function(basegraph, communities.object,vertex.colors,vertex.sizes) {
  plot(basegraph, sub = communities.object$subtitle
    ,mark.groups = communities.object, mark.expand = 2, edge.width=0.2 
         ,edge.color = "lightgrey"
         ,mark.col = communities.object$community.colors
         ,mark.border="black"
         ,vertex.size=vertex.sizes
         ,vertex.color=vertex.colors)
  legend("bottomright",inset=.02,title="Dominant cell type"
         ,fill=communities.object$local.palette
         ,legend=names(communities.object$local.palette) 
         )
}
```

```{r, fig.width = 7, fig.height = 7}
plot.cases.and.controls(g08 ,g08.infomap ,case.control.colors, case.control.size)
```










