---
title: "Plotting Relationship between Physical Distance and Divergence of Cell Type Distribution"
output: md_document
---
# Plotting Relationship between Physical Distance and Divergence of Cell Type Distribution 

```{r, include=FALSE}
library(Seurat)
library(tidyr)
library(ggplot2)
library(igraph)
library(FNN) #KL.divergence
library(philentropy) #JS divergence 'JSD()'
```


The spatial data to be analyzed is available publically from 10x Genomics. 1,072 spatially tagged samples for 47,094 genes. Retreive data:
```{r, include=FALSE}
cortex_final <- readRDS('../2/cortex_final.rds') #342MB
spatial.data <- data.frame(cortex_final[['anterior1']]@coordinates[,c('col','row')]) #see "exploring anchor based label transfer".md
spatial.data['row'] <- spatial.data['row'] * -1
colnames(spatial.data) <- c('x','y')
spatial.data <- cbind(spatial.data,t(as.matrix((cortex_final[['predictions']]@data))))
vals <- as.matrix(spatial.data[,-c(1,2,26)])
dist.matrix <- as.matrix(dist(data.frame(spatial.data[,1:2]))) #see "setting a distance threshold".md```
```

```{r}
head(spatial.data[1:5,1:3])
head(dist.matrix[1:5,1:3])
```

### Using Jensen-Shannon Divergence function from philentropy package
JS Divergence for 2nd and 3rd samples. https://cran.r-project.org/web/packages/philentropy/vignettes/Information_Theory.html
```{r}
suppressMessages(test<-as.numeric(JSD(vals[c(2,3),])))
test
```

JS Divergence for identical distributions is zero:
```{r}
P <- 1:10
Q <- 20:29
JSD(rbind(P,Q))
```

Calculate Jensen-Shannon Divergence for every pair of samples' predicted cell type distributions:
```{r}
n <- nrow(spatial.data)
results <- matrix(data=NA,nrow=n,ncol=n)
for (i in seq(n)){
  for (v in seq(n)){
    if (!is.na(results[v,i])) {
      results[i,v] <- results[v,i]
      next
    }
    current <- 0
    if (i != v) {suppressMessages(current <- as.numeric(JSD(vals[c(i,v),])))}
    results[i,v] <- current
  }
}
dim(results)
sum(is.na(results))
```

```{r}
plot.single.js.vs.distance <- function(target, color.by.cluster=FALSE){
  data <- data.frame(cbind(results[-target,target], dist.matrix[-target,target]))
  colnames(data) <- c("JS.Divergence","Physical.Distance")
  UMAP_cluster <- if (color.by.cluster) {cortex_final$seurat_clusters[-target]}
  ggplot(data,aes(x=JS.Divergence,y=Physical.Distance,col=UMAP_cluster)) +
    ggtitle("Divergence vs. Distance",subtitle=paste("Sample",target,"out of",nrow(results))) +
           geom_point(shape=19, alpha=1/4
                      ,position=position_jitter(width=.01,height=.01)
                      )
}
```

Plot JSD vs. phsyical distance for the first sample:
```{r}
plot.single.js.vs.distance(1)
```

```{r}
plot.single.js.vs.distance(123)
plot.single.js.vs.distance(450)
plot.single.js.vs.distance(450, color.by.cluster=TRUE)
```

```{r}
rws <- seq(nrow(results))
jsd.vs.dist <- cbind(results[rws>1,1], dist.matrix[rws>1,1])
typeof(jsd.vs.dist)
dim(jsd.vs.dist)
colnames(jsd.vs.dist) <- c("JS.Divergence","Physical.Distance")
colnames(jsd.vs.dist)
for (i in seq(2:nrow(results)-1)){
  #print (dim(cbind(results[rws>i,i], dist.matrix[rws>i,i])))
  jsd.vs.dist <- rbind(jsd.vs.dist,cbind(results[rws>i,i], dist.matrix[rws>i,i]))
}
typeof(jsd.vs.dist)
dim(jsd.vs.dist)
jsd.vs.dist <- data.frame(jsd.vs.dist)

```

```{r, fig.width = 12, fig.height = 12}
ggplot(data=jsd.vs.dist,aes(x=JS.Divergence, y=Physical.Distance)) +
           geom_point(shape=19, alpha=.01, size=.5
                      ,position=position_jitter(width=.01,height=.01)
                      )
```


