---
title: "R Notebook"
output: md_document
---

```{r, include=FALSE}
library(Seurat)
library(tidyr)
library(ggplot2)
library(igraph)
library(FNN) #KL.divergence
library(philentropy) #JS divergence 'JSD()'
```


The spatial data to be analyzed is available publically from 10x Genomics. 1,072 spatially tagged samples for 47,094 genes. Retreive data:
```{r, include=FALSE}
cortex_final <- readRDS('cortex_final.rds') #342MB
spatial.data.frame <- data.frame(cortex_final[['anterior1']]@coordinates[,c('col','row')]) #see "exploring anchor based label transfer".md
spatial.data.frame['row'] <- spatial.data.frame['row'] * -1
colnames(spatial.data.frame) <- c('x','y')
spatial.data.frame <- cbind(spatial.data.frame,t(as.matrix((cortex_final[['predictions']]@data))))
expression.vals <- as.matrix(spatial.data.frame[,-c(1,2,26)])
distance.matrix <- as.matrix(dist(data.frame(spatial.data.frame[,1:2]))) #see "setting a distance threshold".md```
```


```{r}
head(spatial.data.frame[1:5,1:3])
head(distance.matrix[1:5,1:3])
```

Calculate Jensen-Shannon Divergence for every pair of samples' predicted cell type distributions:
```{r}
n <- nrow(spatial.data.frame)
jsd.matrix <- matrix(data=NA,nrow=n,ncol=n)
for (i in seq(n)){
  for (v in seq(n)){
    if (!is.na(jsd.matrix[v,i])) {
      jsd.matrix[i,v] <- jsd.matrix[v,i]
      next
    }
    current <- 0
    if (i != v) {suppressMessages(current <- as.numeric(JSD(expression.vals[c(i,v),])))}
    jsd.matrix[i,v] <- current
  }
}
rm(i,v,current)
dim(jsd.matrix)
sum(is.na(jsd.matrix))
```

Save the JSD and the physical distance for every pairwise combo of samples. 1,072 samples -> 575,127 tuples.
```{r}
seq.n <- seq(n)
jsd.vs.dist <- cbind(jsd.matrix[seq.n > 1,1], distance.matrix[seq.n > 1,1]) #load all jsd and distances from 1st sample
#typeof(jsd.vs.dist)
#dim(jsd.vs.dist)
colnames(jsd.vs.dist) <- c("JS.Divergence","Physical.Distance")
colnames(jsd.vs.dist)
for (i in seq(2:n-1)){
  #print (dim(cbind(jsd.matrix[seq.n > i,i], dist.matrix[seq.n>i,i])))
  jsd.vs.dist <- rbind(jsd.vs.dist,cbind(jsd.matrix[seq.n>i,i], distance.matrix[seq.n > i,i]))
}
rm(i)
#typeof(jsd.vs.dist)
#dim(jsd.vs.dist)
jsd.vs.dist <- data.frame(jsd.vs.dist)
summary(jsd.vs.dist)
```

Maybe find a suitable cutoff (.1? .15?) to define "closely matched" pairs. Then draw a graph and link only samples that are closely matched. Then... do shared neighbor algo or community detection??

As we use stricter and stricter cutoffs (down to 0), only identical samples will pass the threshold. So any shared neighbor analysis or even just using those JSD under the threshold as the edges of a graph... will only show the most strongly identical samples. 

Then as we loosen the threshold, more edges should get added. these are the similar but not totally identical samples (maybe indicating some sort of gradient).

This is still just ad hoc making algo to split tissue into subtissues (as a gradient, even, where subtissues collide). When graphing, should be like FEMA flood maps - some places are clearly red, others clearly blue. Then a border area between the two is a blue-and-red diagonal-striped mix of the two.

Some identical sample-pairs may be dispersed (not physically connected). May need to have some special approach to identify and call those out for visualization too! [eg Identical samples, by JSD, but they are separated by more than 3 edges in the physical distance graph -> this is a dispersed type of sample].

**For harmanci: when you work, is it in bursts? Work with collaborators or alone? How well must you know the underlying lab techniques (I have a hard time with all the varied spatial methods)?**


```{r}
quantile(jsd.vs.dist[,1],seq(0,.1,.01)) #percnetiles of all 500k+ JSD values
threshold.percentile <- .0075
threshold.jsd.value <- as.numeric(quantile(jsd.vs.dist[,1],threshold.percentile))
threshold.percentile * length(jsd.vs.dist[,1]) #number of connections past this threshold
threshold.percentile * length(jsd.vs.dist[,1]) / n #avg number of connections per node (sample)
```


```{r}
i <- 1
#length(jsd.matrix[-i,i]) #every sample will have 1,071 others to compare to
boxplot(jsd.matrix[-i,i])
length(which(jsd.matrix[,i] < threshold.jsd.value)) #number of samples with JSD < threshold percentile for sample i
rm(i)

#jsd.matrix[,i][(which(jsd.matrix[,i] < threshold.jsd.value))]
jsd.less.than.threhold <- as.matrix(jsd.matrix < threshold.jsd.value)
jsd.less.than.threhold[1:5,1:5]
```

Plot only edges where JSD is less than the threshold (0.0075%ile of all JSD values). This is the strictest way to join samples: only when JSD matches very well, relative to the whole tissue.
```{r, fig.width = 12, fig.height = 12}
g <- graph_from_adjacency_matrix(jsd.less.than.threhold, mode="undirected", diag = FALSE)
 #iGraph plotting help: https://kateto.net/netscix2016.html
layout.for.igraph <- as.matrix(spatial.data.frame[,1:2])
plot(g, vertex.label=NA, vertex.size=1
     , layout = layout.for.igraph
     , edge.color=adjustcolor("SkyBlue2", alpha.f=.05)) #layout_with_kk is pretty good too
```




# June 19th... try shared neighbor (Jaccard distance)
We already have the JSD for every single sample. What if we calculate Jaccard distance for all pairs of samples, using top 10 smallest JSD (the samples int he top 1% of smallest JSD)?

Calculate Jaccard Distance for every pair of samples, by setting percentile threhold of 0.01 for each sample and comparing overlap:
```{r}
#list(c(1,2,3),c(4,5,6,7))
find.jsd.peers <- function(i, threshold=.01) which(jsd.matrix[,i] <= quantile(jsd.matrix[,i], threshold))
lowest.jsd.peers <- lapply(seq(n),find.jsd.peers)
lowest.jsd.peers[1]
            
            
jaccard.matrix <- matrix(data=NA,nrow=n,ncol=n)
for (i in seq(n)){
  for (v in seq(n)){
    if (!is.na(jaccard.matrix[v,i])) {
      jaccard.matrix[i,v] <- jaccard.matrix[v,i]
      next
    }
    current <- 0
    if (i != v) {#suppressMessages(current <- as.numeric(JSD(expression.vals[c(i,v),])))
      #a <- which(jsd.matrix[,i] <= quantile(jsd.matrix[,i], .01))
      #b <- which(jsd.matrix[,v] <= quantile(jsd.matrix[,v], .01))
      a <- lowest.jsd.peers[i]
      b <- lowest.jsd.peers[v]
      c <- length(union(a,b))
      if (c >0) {
        current <- length(intersect(a,b)) / c
      }
      }
    jaccard.matrix[i,v] <- current
  }
}
rm(i,v,current,a,b,c,find.jsd.peers)
dim(jaccard.matrix)
sum(is.na(jaccard.matrix))
```


Plot those with Jaccard distnace > .01, based on samples' peers with JSD in top 1% for that sample. 
```{r, fig.width = 12, fig.height = 12}
g <- graph_from_adjacency_matrix(as.matrix(jaccard.matrix > .01), mode="undirected", diag = FALSE)
plot(g, vertex.label=NA, vertex.size=1
     , layout = layout.for.igraph
     #, edge.color=adjustcolor("SkyBlue2", alpha.f=.05)
     )
```

Try again, but use a less-restrictive JSD percentile threshold... try 0.10 (instead of 0.01) for selecting "neighbors" to evaluate for Jaccard:
```{r, fig.width = 12, fig.height = 12}
find.jsd.peers <- function(i, threshold=.1) which(jsd.matrix[,i] <= quantile(jsd.matrix[,i], threshold))
lowest.jsd.peers <- lapply(seq(n),find.jsd.peers)
jaccard.matrix <- matrix(data=NA,nrow=n,ncol=n)
for (i in seq(n)){
  for (v in seq(n)){
    if (!is.na(jaccard.matrix[v,i])) {
      jaccard.matrix[i,v] <- jaccard.matrix[v,i]
      next
    }
    current <- 0
    if (i != v) {#suppressMessages(current <- as.numeric(JSD(expression.vals[c(i,v),])))
      #a <- which(jsd.matrix[,i] <= quantile(jsd.matrix[,i], .01))
      #b <- which(jsd.matrix[,v] <= quantile(jsd.matrix[,v], .01))
      a <- lowest.jsd.peers[i]
      b <- lowest.jsd.peers[v]
      c <- length(union(a,b))
      if (c >0) {
        current <- length(intersect(a,b)) / c
      }
      }
    jaccard.matrix[i,v] <- current
  }
}
g <- graph_from_adjacency_matrix(as.matrix(jaccard.matrix > .51), mode="undirected", diag = FALSE)
plot(g, vertex.label=NA, vertex.size=1, layout = layout.for.igraph)
```
Seems like there are 2 large bands of samples that have very siimlar (high JSD) cell type compositions.

I wonder how the cell type distributions look for the samples in between the two major bands are like??

There are some other samples that have one or two  or three closely identical peers, but are physical separated. Again, what is the dist like for those samples in between? (Some of those could be noise or due ot random chance, that the compositions are so similar, but some may really represent dispersed subtissue types(?).) 

