---
title: "R Notebook"
output: md_document
---

For legacy 5330 project, compared group 1 to group 0 (cancer vs non cancer sample
log2( 10^gene means / 10^other gene means )

Then select only genes with absolute value > 2, and sort before passing into the cluster profiler function.
```{r, echo=FALSE}
library(org.Hs.eg.db)
library(clusterProfiler)
library(AnnotationDbi)
```

Load expression values (transformed and normalized, limited to 16k genes from 31k originally.
```{r}
expression.vals <- readRDS('expression.vals.rds')
g4.border.nodes <- readRDS('g4.border.nodes.rds')
g8.border.nodes <- readRDS('g8.border.nodes.rds')
gene.list <- readRDS('gene.list.rds')
SCT <- readRDS('SCT.rds')
```

Cell types data (`expressoin.vals`) holds composition of 23 cell types for all 1,072 spots. Transformed counts data (`SCT`) has normalized gene counts of 16,017 genes for every spot.
```{r}
n <- dim(expression.vals)[1]
dim(expression.vals)
dim(SCT)
```

Split the 1,072 samples into two groups "g0" and "g1". Log10 transform (this was done for 5330 project).
```{r}
print (c('border samples:', length(g4.border.nodes)))
print (c('non border samples:', n-length(g4.border.nodes)))
```

Take gene means for each of the 2 different groups of spots, and plot the log-transformed ratio.
```{r}
g0.means <- rowMeans(SCT[,-g4.border.nodes])
g1.means <- rowMeans(SCT[,g4.border.nodes])
#create gene list
gene.list.1.0 <- 10^g1.means / 10^g0.means #raw values are all log10 transormed. undo that.
gene.list.1.0 <- log(gene.list.1.0, base=2)
plot(sort(gene.list.1.0))
abline(0,0)
```

Use bioconductor package to convert gene names into Entrez ids. Remove duplicate Entrez located for 3 of the genes:
```{r}
annot <- AnnotationDbi::select(org.Hs.eg.db, keys=toupper(gene.list), columns=c("SYMBOL","ENTREZID")
                , keytype="SYMBOL", multiVals="first")
which(table((annot[,1]))>1)
length(unique(annot[,1]))
sum(annot[,1] %in% c("MEMO1","MMD2","TEC")) #6 entrez ids for these 3 gene symbols
#which(annot[,1] %in% c("MEMO1","MMD2","TEC"))
annot[which(annot[,1] %in% c("MEMO1","MMD2","TEC")),]
```
Remove the duplicate rows, to stay at 16,017 gene rows.
```{r}
annot <- annot[-c(4848,5443,14992),]
head(annot)
```

Alternate plot of log-ratios for all 16k genes:
```{r}
hist((gene.list.1.0))
```
Save differentiall expressed genes. Limit to only genes that mapped to an Entrez ID & had log-ratio > 0.3
```{r}
differential.genes <- annot[!is.na(annot[,2]) & abs(gene.list.1.0) > 0.3,2]
differential.genes
write.table(differential.genes, file="differential_entrez_logratio_abs_greater_point3.txt"
            , row.names = FALSE, col.names = FALSE, quote = FALSE)
```

save separate lists of differentially up and down expressed genes:
```{r}
differential.genes <- annot[!is.na(annot[,2]) & gene.list.1.0 > 0.3,2]
length(differential.genes)
write.table(differential.genes, file="differential_entrez_logratio_greater_point3.txt"
            , row.names = FALSE, col.names = FALSE, quote = FALSE)

differential.genes <- annot[!is.na(annot[,2]) & gene.list.1.0 < -0.3,2]
length(differential.genes)
write.table(differential.genes, file="differential_entrez_logratio_less_point3.txt"
            , row.names = FALSE, col.names = FALSE, quote = FALSE)
```
### Differential genes (n=64)
```{r}
ggo <- groupGO(gene     = differential.genes,
               OrgDb    = org.Hs.eg.db,
               ont      = "MF",
               level    = 2,
               readable = TRUE)
ggo@result = ggo@result[order(ggo@result[,"Count"],decreasing=TRUE),]
ggo@result

barplot(ggo,showCategory=8)
```

### Up expressed genes
```{r}
differential.genes <- annot[!is.na(annot[,2]) & gene.list.1.0 > 0.3,2]
ggo <- groupGO(gene     = differential.genes,
               OrgDb    = org.Hs.eg.db,
               ont      = "MF",
               level    = 2,
               readable = TRUE)
ggo@result = ggo@result[order(ggo@result[,"Count"],decreasing=TRUE),]
ggo@result
```


### Down expressed genes
```{r}
differential.genes <- annot[!is.na(annot[,2]) & gene.list.1.0 < -0.3,2]
ggo <- groupGO(gene     = differential.genes,
               OrgDb    = org.Hs.eg.db,
               ont      = "MF",
               level    = 3,
               readable = TRUE)
ggo@result = ggo@result[order(ggo@result[,"Count"],decreasing=TRUE),]
ggo@result
```



```{r}
differential.genes <- annot[!is.na(annot[,2]) & gene.list.1.0 > 0.3,2]
ggo <- groupGO(gene     = differential.genes,
               OrgDb    = org.Hs.eg.db,
               ont      = "MF",
               level    = 3,
               readable = TRUE)
ggo@result = ggo@result[order(ggo@result[,"Count"],decreasing=TRUE),]
ggo@result
```


```{r}
differential.genes
```


```{r}
length(gene.list.1.0)
length(gene.list)
annot['ENTREZID']
test <- gene.list.1.0
test[1:5]
names(test)[1:5]
typeof(unlist(annot['ENTREZID']))
names(test) <- unlist(annot['ENTREZID'])
test <- sort(test, decreasing=TRUE)
test[1:5]
```



# Older Spring 2019 code below.
```{r}
gene <- names(test)[abs(test) > 2]
ggo <- groupGO(gene     = gene,
               OrgDb    = org.Hs.eg.db,
               ont      = "MF",
               level    = 2,
               readable = TRUE)
#order by count
ggo@result = ggo@result[order(ggo@result[,"Count"],decreasing=TRUE),]
ggo@result
```

```{r}
barplot(ggo,showCategory=8)
```



Further information about enirchment scores:
https://yulab-smu.github.io/clusterProfiler-book/chapter2.html#gene-set-enrichment-analysis

```{r}
#KEGG Gene Set Enrichment Analysis
kk2 <- gseKEGG(geneList     = test,
               organism     = 'hsa',
               nPerm        = 1000,
               minGSSize    = 100,
               pvalueCutoff = 0.05,
               verbose      = FALSE)
kk2@result = kk2@result[order(kk2@result[,"setSize"],decreasing=TRUE),]
kk2@result
```

```{r}
#KEGG over representation Analysis
kk3 <- enrichKEGG(gene    = names(test),
               organism     = 'hsa',
               pvalueCutoff = 0.05
  )
kk3@result[,c("Description","GeneRatio","p.adjust","Count")]
```


```{r}
#plot(kk2@result[,"setSize"])
```



```{r}
#library(GSEABase)
#library(tidyr)
```

```{r}
#gene <- names(test)[abs(test) > 2]
#wpgmtfile <- system.file("extdata", "wikipathways-20180810-gmt-Homo_sapiens.gmt", package="clusterProfiler")
#wp2gene <- read.gmt(wpgmtfile) #i had to install GSEAbase first ??
#wp2gene <- wp2gene %>% tidyr::separate(ont, c("name","version","wpid","org"), "%")
#wpid2gene <- wp2gene %>% dplyr::select(wpid, gene) #TERM2GENE
#wpid2name <- wp2gene %>% dplyr::select(wpid, name) #TERM2NAME
#ewp <- enricher(gene, TERM2GENE = wpid2gene, TERM2NAME = wpid2name)
#ewp <- setReadable(ewp, org.Hs.eg.db, keyType = "ENTREZID")
#ewp@result = ewp@result[order(ewp@result[,"Count"],decreasing=TRUE),]
#ewp@result
```

```{r}
#barplot(ewp,showCategory=12)
```


```{r}
#ewp2 <- GSEA(test, TERM2GENE = wpid2gene, TERM2NAME = wpid2name, verbose=FALSE)
#ewp2 <- setReadable(ewp2, org.Hs.eg.db, keyType = "ENTREZID")
#ewp2@result = ewp2@result[order(ewp2@result[,"setSize"],decreasing=TRUE),]
#ewp2@result
```

```{r}
#gset@featureData@data[which(gset@featureData@data[,"Gene.ID"]==100128545),c("ID","Gene.ID")]
```
