---
title: "Try DESEQ2"
output: html_notebook
---



```{r, echo=FALSE}
library(org.Hs.eg.db)
#library(clusterProfiler)
library(AnnotationDbi)
```

Load expression values (transformed and normalized, limited to 16k genes from 31k originally.

Cell types data (`expressoin.vals`) holds composition of 23 cell types for all 1,072 spots. Transformed counts data (`SCT`) has normalized gene counts of 16,017 genes for every spot.
```{r}
expression.vals <- readRDS('expression.vals.rds')
gene.list <- readRDS('gene.list.rds')
SCT <- readRDS('SCT.rds')
Spatial <- readRDS('cortex_Spatial.rds')

n <- dim(expression.vals)[1]
dim(expression.vals)
dim(SCT)
```



```{r}
g0 <- "64 68 176 191 248 264 286 305 334 340 344 420 445 494 570 675 710 760 774 780 790 849 929 956 964 971 987 1004 1009 1065"
g0 <- as.numeric(unlist(strsplit(g0," ")))
#"cases" or "treated", go here:
g1 <- "12 33 35 38 54 121 146 189 304 306 352 395 409 484 492 574 609 707 726 828 1011 490 462 406 1031 1037"
g1 <- as.numeric(unlist(strsplit(g1," ")))
g0
```
### Need to be using raw counts, instead of transformed SCT counts!
```{r}
cts <- Spatial[,c(g0,g1)]
dim(cts)
head(cts[,1:10])
coldata <- as.matrix(as.factor(c(rep('untreated',length(g0)),rep('treated',length(g1)))))
rownames(coldata) <- colnames(cts)
colnames(coldata) <- 'condition'
coldata
#coldata
```


```{r}
library("DESeq2")
```

```{r}
dds <- DESeqDataSetFromMatrix(countData = cts,
                              colData = coldata,
                              design = ~ condition)
dds$condition <- relevel(dds$condition, ref = "untreated")
dds
```

Can add Entrez IDs into mcol(dds), as additional metadata




Prefiltering:
```{r}
keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep,]
```

Differential Expression Analysis
```{r}
dds <- DESeq(dds)
res <- results(dds)
res
```

```{r}
plot(-log(sort(res$padj)))
abline(-log(.05),0)
```


```{r}
rownames(res)[which(res$padj<.05)]
```

```{r}
plotCounts(dds, gene=which.min(res$padj), intgroup="condition")
```




