---
title: "R Notebook"
output: md_document
---
Will need to rerun for a slideseq dataset too! Seurat vignettes updated 7-16-2020.
 
```{r, include=FALSE}
library(Seurat)
library(tidyr)
library(ggplot2)
library(igraph)
library(FNN) #KL.divergence
library(philentropy) #JS divergence 'JSD()'
library(RColorBrewer)
```

The spatial data to be analyzed is available publicly from 10x Genomics. 1,072 spatially tagged samples for 47,094 genes. Used Seurat anchor-based algorithm to label each sample spot with composition of 23 different cell types:
```{r, echo=FALSE}
cortex_final <- readRDS('../cortex_final.rds') #342MB
spatial.data.frame <- data.frame(cortex_final[['anterior1']]@coordinates[,c('col','row')]) #see "exploring anchor based label transfer".md
spatial.data.frame['row'] <- spatial.data.frame['row'] * -1
colnames(spatial.data.frame) <- c('x','y')
spatial.data.frame <- cbind(spatial.data.frame,t(as.matrix((cortex_final[['predictions']]@data))))
expression.vals <- as.matrix(spatial.data.frame[,-c(1,2,26)])
distance.matrix <- as.matrix(dist(data.frame(spatial.data.frame[,1:2]))) #see "setting a distance threshold".md
distance.matrix <- apply(ifelse(distance.matrix==0, NA, distance.matrix),1,sort.list) #Rank samples on proximity to one another. Do not count if distance=0 (the diagonal of the distance matrix)
head(spatial.data.frame)
```

Plot original stained tissue image and UMAP clusters.
```{r}
SpatialDimPlot(cortex_final, crop = FALSE, pt.size.factor = 0, label.size = 3)
SpatialDimPlot(cortex_final, crop = TRUE, pt.size.factor = 0, label.size = 3)
SpatialDimPlot(cortex_final, crop = TRUE, label = FALSE)
rm(cortex_final) # <- uncomment this to save RAM
```


Save X-Y coordinates for plotting network graphs in the same shape as the original slide:
```{r}
layout.for.igraph <- as.matrix(spatial.data.frame[,1:2]) 
set_default_igraph_attr <- function(g,max.out.degree=NA){
  graph_attr(g,"layout") <- layout.for.igraph
  vertex_attr(g,"size") <- 1 #set here, instead of passing in "vertex.size=1" to plot(g)
  vertex_attr(g,"label") <- NA 
  edge_attr(g,"color") <- adjustcolor("SkyBlue2", alpha.f=1)
  if (is.weighted(g)){
    edge_attr(g, "width") <- E(g)$weight*5
  }
  if (!is.na(max.out.degree)){
    graph_attr(g,"max.out.degree") <- max.out.degree
    graph_attr(g,"main") <- paste("Nodes connected to nearest",max.out.degree,"neighbors")
    if (is.weighted(g)) {graph_attr(g,"main") <- paste(graph_attr(g,"main"),", weighted with JSD")}
  }
  set_default_igraph_attr <- g
}
```

Calculate Jensen-Shannon Divergence for every pair of samples' predicted cell type distributions:
```{r}
n <- nrow(spatial.data.frame)
jsd.matrix <- matrix(data=NA,nrow=n,ncol=n)
for (i in seq(n)){
  for (v in seq(n)){
    if (!is.na(jsd.matrix[v,i])) {
      jsd.matrix[i,v] <- jsd.matrix[v,i]
      next
    }
    current <- 0
    if (i != v) {suppressMessages(current <- as.numeric(JSD(expression.vals[c(i,v),])))}
    jsd.matrix[i,v] <- current
  }
}
rm(i,v,current)
dim(jsd.matrix)
```

New July 16: use distance ranks to create network graph, instead of setting a distance threshold. Build edgelist of closest N neighbors for each node.
```{r}
make_graph_of_closest_neighbors <- function(max.out.degree) {
  #out.degree = max number of edges for each node 
  temp <- distance.matrix[1:max.out.degree,]
  results <- vector('numeric')
  for (i in seq(n)){
    #print (paste("i",i))
    x <- temp[, i]
    #print (x)
    for (v in x){
      #print (paste(".v",v))
      #print (temp[,v])
      if (i %in% temp[, v]){
        #print (paste('yes',i,v))
        results <- c(results,i,v)}
    }
  }
  results <- matrix(results, ncol=2, byrow=TRUE) #build an edge list
  results <- unique(t(apply(results,1,sort))) #after sorting cols 1 & 2, can prevent duplicate rows
  g <- graph_from_edgelist(results, directed=FALSE)
  g <- set_default_igraph_attr(g,max.out.degree=max.out.degree)
  make_graph_of_closest_neighbors <- g
}
```

## Use new function to create networks of closest 4, 6, 8, 12 neighbors:
```{r}
g04 <- make_graph_of_closest_neighbors(4)
g06 <- make_graph_of_closest_neighbors(6)
g08 <- make_graph_of_closest_neighbors(8)
g12 <- make_graph_of_closest_neighbors(12)
summary(g04,graph.attributes=TRUE)
summary(g08,graph.attributes=TRUE)
```

```{r, fig.width = 6, fig.height = 6}
plot(g04)
plot(g08) 
```

## Use JSD to create weighted networks
Weight the networks using 1 - JSD. (JSD=0 for perfectly identical cell type distributions).
```{r}
make_graph_weighted_with_JSD <- function(base_graph){
  #starting with base_graph, weight edges with the Jensen Shannon Divergence of the 2 nodes attached by the edge
  g <- as_adjacency_matrix(base_graph,type="both",sparse=FALSE) > 0 #create adjacency matrix, TRUE or . (null)
  #jsd=0 for identical dist. so use 1 minus JSD and ignore JSD=1 (total dissimilarity)
  g <- ifelse(g & 1-jsd.matrix>0, 1-jsd.matrix, 0) 
  g <- graph_from_adjacency_matrix(g, mode="undirected",weighted=TRUE)
  g <- set_default_igraph_attr(g, max.out.degree = if (!is.null(graph_attr(base_graph,"max.out.degree"))) {graph_attr(base_graph,"max.out.degree")} else {NA})
}
```

```{r, fig.width = 9, fig.height = 9}
g04.weighted <- make_graph_weighted_with_JSD(g04)
g06.weighted <- make_graph_weighted_with_JSD(g06)
g08.weighted <- make_graph_weighted_with_JSD(g08)
g12.weighted <- make_graph_weighted_with_JSD(g12)
plot(g04.weighted)
plot(g08.weighted)
```

## Apply Louvain Community detection
```{r, include=FALSE}
#https://www.r-bloggers.com/community-detection-with-louvain-and-infomap/
#lc04 <- cluster_louvain(g04.weighted) #must remove zero-weight edges first
#lc06 <- cluster_louvain(g06.weighted)
#lc08 <- cluster_louvain(g08.weighted)
#lc12 <- cluster_louvain(g12.weighted)
#when plotting, use "mark." to control the community polygons in the plot
#plot(lc04, g04.weighted, mark.expand = 2, edge.width=1)
#plot(lc08, g08.weighted, mark.expand = 2, edge.width=1)
#identical to: 
#plot(g04.weighted, mark.groups=lc04, mark.expand = 2, edge.width=1 
# ,edge.color=ifelse(crossing(lc04,g04.weighted),"red","black"))
```

## Color By Dominant Cell Type for Each Community
```{r, fig.width = 7, fig.height = 7}
plot_communities_by_dominant_cell_type <- function(basegraph, communities.algo=NA
                                                   , print.table=FALSE, expr=expression.vals){
  #communities.algo is a string: cluster_louvain, cluster_infomap
  communities.object <- NA
  if (is.na(communities.algo) | communities.algo=="cluster_louvain") {
    communities.object <- cluster_louvain(basegraph)
  }else if (communities.algo=="cluster_infomap"){
    communities.object <- cluster_infomap(basegraph)
  }
  if (is.na(communities.object)) {break}
  comm.top.celltypes <- lapply(communities(communities.object),function(z){
    x <- colMeans(expr[z,])
    x <- x[sort.list(x,decreasing = TRUE)]
    if (length(which(x>.3))>0){
      x[which(x>.3)]
    } else if (length(which(x>.15))>0){
      x[which(x>.15)]
    } else {
      x[1:2]
    }
  })
  comm.highest.label <- as.factor(sapply(comm.top.celltypes, function(x){names(x)[1]}))
  comm.labels <- sapply(comm.top.celltypes, function(x){paste(names(x),sprintf(x,fmt = '[%#.2f]'),collapse="; ")})
  names(comm.labels) <- paste("Community #",names(comm.labels),sep="")
  #comm.highest.val <- sapply(comm.top.celltypes, function(x){x[1]},USE.NAMES = FALSE)
  if (print.table) {
    print ("")
    print (paste("Highest Cell Types per Community,",graph_attr(basegraph,"main")))
    print (comm.labels)
  }
  #take care here to use a global to set palette and ensure it is large enough to handle 23 cell types:
  #by using global palette, cell types will retain same color on different plots
  #set alpha to < 1.0, so that community overlaps can be seen
  global.pal <- topo.colors(length(unique(colnames(expression.vals))),alpha=.2)
  names(global.pal) <- colnames(expression.vals)
  local.pal <- global.pal[as.character(unique(comm.highest.label))] #unique colorset used for this instance 
  comm.pal <- local.pal[as.character(comm.highest.label)] #colors used by each community
  plot(basegraph
       #, sub="asdfasdf"
       ,mark.groups=communities.object, mark.expand = 2, edge.width=0.2 
       ,edge.color="lightgrey"
       ,mark.col=comm.pal
       ,mark.border="black")
  legend("bottomright",inset=.02,title="Dominant cell type"
         ,fill=local.pal
         ,legend=names(local.pal) 
         )
}
```


```{r, fig.width = 7, fig.height = 7}
plot_communities_by_dominant_cell_type(g04.weighted, "cluster_louvain", print.table=FALSE)
plot_communities_by_dominant_cell_type(g04.weighted, "cluster_infomap", print.table=FALSE)
```
**^Why is the subtitle still not working? Should print what community detection algo was used, and meta measures like Modularity**

```{r, fig.width = 7, fig.height = 7}
plot_communities_by_dominant_cell_type(g08.weighted, "cluster_louvain")
plot_communities_by_dominant_cell_type(g08.weighted, "cluster_infomap")
```

## Characterize a specific community
```{r, include=FALSE}
lc04 <- cluster_louvain(g04.weighted)
#Explore communities igraph objects:
boxplot(as.numeric(lapply(communities(lc04),length)),main="Size of communities, network of nearest 4 neighbors")
length(lc04)
sizes(lc04)
modularity(lc04)
algorithm(lc04)
is_hierarchical(lc04)
length(crossing(lc04,g04.weighted)) #says if each edge is a cross of two communities
length(membership(lc04)) #says which group each vertex is in

#"compare" function is interesting...
g <- make_graph("Zachary")
sg <- cluster_spinglass(g)
le <- cluster_leading_eigen(g)
compare(sg, le, method="rand")
compare(membership(sg), membership(le))
rm(g,sg,le)
```

Create functions for exploring specific communities:
**Broken: wont print x axis if only one cell type in boxplot**
```{r}
plot_community_location <- function(basegraph,communities.object,community.i){
  #plot a single community: 
  plot(basegraph, mark.groups=communities.object[[community.i]], mark.expand = 2, edge.width=1
       ,edge.color=ifelse(crossing(communities.object,basegraph),"red","black")
       #,main=paste(graph_attr(basegraph,"main"),paste(", community #",community.i,sep=""))
       ,sub=paste("Community #",community.i," highlighted (n=",length(communities.object[[community.i]]),")",sep="")
       )
}

get_prevalent_cell_types <- function(communities.object, community.i, cutoff=.05, expr=expression.vals){
  targets <- communities.object[[community.i]]
  x <- expr[targets,]
  row.names(x) <- as.character(targets)
  prevalent_cell_types <- names(which(apply(x,2,mean)>cutoff)) #cell types that average over 5% of composition of samples in this community
  x[,prevalent_cell_types]
}

plot_prevalent_cell_types <- function(communities.object, community.i, cutoff=.05, expr=expression.vals){
  title <- paste("Prevalent Cell Types for Community #",community.i," (n=",sep="")
  title <- paste(title,length(communities.object[[community.i]]),")",sep="")
  boxplot(get_prevalent_cell_types(communities.object, community.i, cutoff, expr),main=title)
}

plot_cell_type_distributions <- function(communities.object, community.i, cutoff=.05, expr=expression.vals){
  y <- as.data.frame(get_prevalent_cell_types(communities.object, community.i))
  barplot(t(y), col=as.factor(colnames(y)), legend.text = TRUE
        ,names.arg=rep("",dim(y)[1])
        ,main=paste('Cell Type Compositions for Samples in Community #',1,sep="")
        )
}

plot_cell_type_distributions_gg <- function(communities.object, community.i, cutoff=.05, expr=expression.vals){
  #cant use ggplot when doing multiple plots together, but it would be nice...
  y <- as.data.frame(get_prevalent_cell_types(communities.object, community.i))
  y['sample'] <- as.character(row.names(y))
  y <- y %>% pivot_longer(-sample,names_to='cell.type',values_to='composition')
  ggplot(y, aes(fill=cell.type, y = composition, x = sample))+
    ggtitle(paste('Cell Type Compositions for Samples in Community #',community.i,sep=""))+
    geom_bar(position="stack", stat="identity")+
    theme(axis.text.x=element_blank())
}

plot_EDA_for_community <- function(basegraph,community.i,communities.object=NA, cutoff=.05, expr=expression.vals){
  #par(mfrow=c(2,1))
  if (is.na(communities.object)) {communities.object <- cluster_louvain(basegraph)}
  plot_community_location(basegraph,communities.object,community.i)
  plot_cell_type_distributions(communities.object,community.i,cutoff,expr)
  plot_prevalent_cell_types(communities.object,community.i,cutoff,expr)
  #p1 + p2 #cant use mfrow for ggplot (plot_prevalent_cell_types)
}
```


The 65 nodes in community #1 seem to have similar cell type distributions:
```{r}
base <- g04.weighted
i <- 1
plot_EDA_for_community(base, i)
```

```{r}
base <- g08.weighted
i <- 13
plot_EDA_for_community(base, i)
```


# Create random graphs, to assess frequency of getting communities of similar size
Can see if getting large community membership size is probable. Will use both Modularity and mean community size as the measure (high modularity is interpreted as dense intra community connections and sparse inter community connections). 
```{r}
deg <- 4
B <- 10000

simulation.results <- t(replicate(B,{
null.g <- sample_degseq(rep(deg,n), method = "simple.no.multiple")
null.g <- set_default_igraph_attr(null.g)
null.cl <- cluster_louvain(null.g)
c(modularity(null.cl),mean(membership(null.cl)))
}))
colnames(simulation.results) <- c("modularity","mean community size")
```
Repeated 1,000 times, modularity for random 4-degree networks is no where close to what is observed in the tissue:
```{r}
boxplot(simulation.results[,1], ylim=c(.5,1)
        , main = paste("Modularity:",B,"simulations vs. Observed (red)")
        , sub = get.graph.attribute(g04.weighted,"main"))
points(modularity(lc04),col="red",pch=19)
```

After 1,000 runs, very rare for mean community size to reach what is observed in the actual Louvain(4) network.
```{r}
boxplot(simulation.results[,2]
        , main = paste("Mean community size:",B,"simulations vs. Observed (red)")
        , sub = get.graph.attribute(g04.weighted,"main"))
points(mean(membership(lc04)),col="red",pch=19)
```

### Later
* Interactive Shiny app - use sliders to adjust number of nearest neighbor edges, and to "blob" together communities, * Use radio buttons to select which community detection algo to use.
* Hover over community to see its name and a chart with aggregate cell type distribution of that community
* Apply algorithm to slide-seq data, and another 10x dataset (need to find anchoring reference though)




# Community Annotations:
When "blobbing" together communities, may need to draw polygons individually, instead of relying on plot(). Or would be nice to label communities with their dominant cell type.
https://stackoverflow.com/questions/47971179/igrph-adding-text-to-community-plot
```
for (g in seq_along(mark.groups)) {
v <- V(graph)[mark.groups[[g]]]
igraph.polygon(layout[v, , drop = FALSE], vertex.size = vs, 
               expand.by = mark.expand[g]/200, shape = mark.shape[g], 
               col = mark.col[g], border = mark.border[g])
```

Manually adding text:
`text(c(-1.15, 0.8, 0.9), c(0.35, -0.7, 0.8), c("A", "B", "C"))`