---
title: "Setting a distance threshold for conversion from Spatial Tx into a Network Graph"
output: md_document
---
# Setting a distance threshold for conversion from Spatial Tx into a Network Graph
May 26th, 2020

```{r, include=FALSE}
library(Seurat)
library(tidyr)
library(ggplot2)
library(igraph)
```


The spatial data is available publically from 10x Genomics. 1,072 spatially tagged samples for 47,094 genes.
```{r, include=FALSE}
cortex_final <- readRDS('../2/cortex_final.rds') #342MB
```
```{r}
cortex_final
```


```{r}
spatial.data <- data.frame(cortex_final[['anterior1']]@coordinates[,c('col','row')])
spatial.data['row'] <- spatial.data['row'] * -1 #final row of cell type predictions is a "max" that can be removed
colnames(spatial.data) <- c('x','y')
spatial.data <- cbind(spatial.data,t(as.matrix((cortex_final[['predictions']]@data))))
dim(spatial.data)
spatial.data[1:10,1:4]
```

```{r}
dist.matrix <- as.matrix(dist(data.frame(spatial.data[,1:2])))
n <- dim(dist.matrix)[1]
dim(dist.matrix)
dist.matrix[1:10,1:3]
```

Median distance is 29.15, mean is 33.58, the minimum distance between two samples is 1.4, while the 2nd percentile distance is 4.0
```{r}
dist.matrix.no.diag <- dist.matrix
for (i in seq(n)) {
  dist.matrix.no.diag[i,i] <- NA
}
summary(as.numeric(dist.matrix.no.diag))
search.space <- quantile(dist.matrix.no.diag, probs = c(0,.02), na.rm = TRUE)
search.space
```

By testing different threholds, we can see that 86% of samples will have four neighbors when using a 1.9 distance threshold, while 79% will have eight neighbors when using a 2.1 distance threshold.
```{r}
cutoffs <- seq(search.space[1], search.space[2], length.out = 20) #50 bins between the minimum and the 2nd percentile distances
getmode <- function(v) {
   uniqv <- unique(v)
   max(uniqv[which.max(tabulate(match(v, uniqv)))])
}
final.results <- data.frame(threshold=double()
                            ,median.neighbors=double()
                            ,mode.neighbors=double()
                            ,mode.fraction=double()
                            ,neighborless=integer()
                            )
for (threshold in cutoffs){
  x <- as.numeric(colSums(dist.matrix > 0 & dist.matrix < threshold))
  median.neighbors <- median(x)
  mode.neighbors <- getmode(x)
  mode.fraction <- sum(x==mode.neighbors) / length(x)
  neighborless <- sum(x==0)
  final.results <- rbind(final.results,data.frame(threshold,median.neighbors,mode.neighbors,mode.fraction,neighborless))
}
final.results[1:20,]
```

We have a range of thresholds to choose an output network where the majority of nodes have either 4, 8, 12, or 20 neighbors:
```{r}
ggplot(final.results, aes(x=threshold, y=mode.fraction, fill=as.factor(mode.neighbors))) + 
  geom_area() + 
  scale_x_continuous('threshold',breaks=seq(search.space[1], search.space[2], length.out = 5))
```

Make a function that uses a distance threshold to create an undirected graph. Can later work on an extension that would weight and/or direct the edges in the output network too.
```{r}
make_graph <- function(threshold) {
   graph_from_adjacency_matrix(as.matrix(dist.matrix > 0 & dist.matrix < threshold), mode="undirected")
}
l <- as.matrix(spatial.data[,1:2]) #use layout coordinates from original data
```

Select a random target sample to plot:
```{r}
plot_random_node <- function(threshold,seed=1989) {
  g <- make_graph(threshold)
  set.seed(seed)
  target.sample <- sample(rownames(dist.matrix),1)
  inc.edges <- incident(g,  V(g)[target.sample], mode="all")
  ecol <- rep("gray80", ecount(g))
  ecol[inc.edges] <- "orange"
  
  vcol <- rep("grey40", vcount(g)) #color vertices, default is grey40
  vcol[V(g)==target.sample] <- "gold" #neighbors of target are gold
  neigh.nodes <- neighbors(g, V(g)[target.sample], mode="out")
  vcol[neigh.nodes] <- "#ff9d00"
  #grep("red", colors(), value=T)
  vcol[sum(V(g)[target.sample])] <- "red2"
  
  plot(g, vertex.label=NA, vertex.size=2, layout = l
       ,vertex.color=vcol, edge.color=ecol) #layout_with_kk is pretty good too
  title(paste("Neighbors for", target.sample," [distance threshold =",threshold,"]"),cex.main=1,col.main="green")
}
```

Here we can see the effect the maxmium neighbor distance threhold has on the final output graph:
```{r, fig.width = 12, fig.height = 12}
par(mfrow=c(2,2))
plot_random_node(1.9)
plot_random_node(2.1)
plot_random_node(3.1)
plot_random_node(3.4)
```

