---
title: "Spatial Tx Graph Weighted by Correlation of Cell Type Compositions"
output: md_document
---
# Spatial Tx Graph Weighted by Correlation of Cell Type Compositions

```{r, include=FALSE}
library(Seurat)
library(tidyr)
library(ggplot2)
library(igraph)
```

Graph edges can be weighted, see: https://stackoverflow.com/questions/54232821/how-igraph-handle-weights
```{r}
g <- make_ring(10) %>%
  set_vertex_attr("color", value = "red") %>%
  set_vertex_attr("label", value = letters[1:10])
#g
#V(g)
#vertex_attr(g)
E(g)$weight <- seq(1,10,1)
#E(g)
#edge_attr(g)
#strength(g) #total of outgoing weights
plot(g, edge.width=(E(g)$weight))
```



The spatial data to be analyzed is available publically from 10x Genomics. 1,072 spatially tagged samples for 47,094 genes. Retreive data:
```{r, include=FALSE}
cortex_final <- readRDS('../2/cortex_final.rds') #342MB
spatial.data <- data.frame(cortex_final[['anterior1']]@coordinates[,c('col','row')]) #see "exploring anchor based label transfer".md
spatial.data['row'] <- spatial.data['row'] * -1
colnames(spatial.data) <- c('x','y')
spatial.data <- cbind(spatial.data,t(as.matrix((cortex_final[['predictions']]@data))))
vals <- as.matrix(spatial.data[,-c(1,2)])
dist.matrix <- as.matrix(dist(data.frame(spatial.data[,1:2]))) #see "setting a distance threshold".md```
```

Make a function that uses a distance threshold to create an undirected graph. Can later work on an extension that would weight and/or direct the edges in the output network too.
```{r}
make_graph <- function(threshold) {
   g <- graph_from_adjacency_matrix(as.matrix(dist.matrix > 0 & dist.matrix < threshold), mode="undirected")
   graph_attr(g,"threshold") <- threshold
   graph_attr(g,"layout") <- as.matrix(spatial.data[,1:2])
   g
}
```

Select a random target sample to plot:
```{r}
plot_random_node <- function(g, seed=1989) {
  set.seed(seed)
  target.sample <- sample(rownames(dist.matrix),1)
  inc.edges <- incident(g,  V(g)[target.sample], mode="all")
  ecol <- rep("gray80", ecount(g))
  ecol[inc.edges] <- "orange"
  
  vcol <- rep("grey40", vcount(g)) #color vertices, default is grey40
  vcol[V(g)==target.sample] <- "gold" #neighbors of target are gold
  neigh.nodes <- neighbors(g, V(g)[target.sample], mode="out")
  vcol[neigh.nodes] <- "#ff9d00"
  #grep("red", colors(), value=T)
  vcol[sum(V(g)[target.sample])] <- "red2"
  
  plot(g, vertex.label=NA, vertex.size=2#, layout = g$layout #setting a "layout" attribute in graph will make it the default
       ,vertex.color=vcol, edge.color=ecol, edge.width=3) #layout_with_kk is pretty good too
  title(paste("Neighbors for", target.sample," [distance threshold =",g$threshold,"]"),cex.main=1,col.main="green")
}
```

Make graph by linking nodes within "1.9" distance of one another:
```{r, fig.width = 6, fig.height = 6}
g <- make_graph(1.9)
plot_random_node(g)
```

Loop over all edges and delete edges where nodes do not have a high correlation:
https://stackoverflow.com/questions/49834827/louvain-community-detection-in-r-using-igraph-format-of-edges-and-vertices
```{r}
apply_correlation_filter <- function(g, cor.cutoff){
  neighbor.cors <- c()
  for (i in seq(length(E(g)))){
    neighbor.cors <- c(neighbor.cors,cor(vals[names(tail_of(g,E(g))[i]),], vals[names(head_of(g,E(g))[i]),]))
  }
  neighbor.cors[1:5]
  length(neighbor.cors)
  targets <- which(neighbor.cors < cor.cutoff)
  delete_edges(g, targets)
}

g7 <- apply_correlation_filter(g, .7)
g8 <- apply_correlation_filter(g, .8)
g9 <- apply_correlation_filter(g, .9)
```

### Note the effect of keeping only edges where neighbor nodes correlate > .7, > .8 and > .9 
```{r, fig.width = 12, fig.height = 12}
par(mfrow=c(2,2))
seed = 11
plot_random_node(g, seed)
plot_random_node(g7, seed)
plot_random_node(g8, seed)
plot_random_node(g9, seed)
```

Should try community detection. Can't do cliques (complete connected componets). See: https://kateto.net/networks-r-igraph 
or https://igraph.org/c/doc/igraph-Community.html
```{r, fig.width = 12, fig.height = 12}
ceb <- cluster_edge_betweenness(g9)  #makes 500 communities
#hist(membership(ceb))
plot(ceb, g9, vertex.label=NA, vertex.size=2)#, layout = g$layout #setting a "layout" attribute in graph will make it the default
```



```{r, fig.width = 12, fig.height = 12}
clp <- cluster_label_prop(g9)
plot(clp, g9, vertex.label=NA, vertex.size=2)
```

```{r, fig.width = 12, fig.height = 12}
cfg <- cluster_fast_greedy(as.undirected(g9))
#plot(cfg, g9, vertex.label=NA, vertex.size=2, vertex.col = cut_at(cfg, step=100))
```

```{r}
cfg
length(sizes(cfg)) #499
sizes(cfg)
merges(cfg)
cfg[1]
cfg[400]
sapply(cfg,length)
cut_at(cfg, step=250)
#igraph_community_to_membership(cfg)
```


from source: https://www.sixhat.net/finding-communities-in-networks-with-r-and-igraph.html
```{r}
 
# let's see if we have communities here using the 
# Grivan-Newman algorithm
# 1st we calculate the edge betweenness, merges, etc...
ebc <- edge.betweenness.community(g9, directed=F)
5+5
# Now we have the merges/splits and we need to calculate the modularity
# for each merge for this we'll use a function that for each edge
# removed will create a second graph, check for its membership and use
# that membership to calculate the modularity
mods <- sapply(0:ecount(g9), function(i){
  g2 <- delete.edges(g9, ebc$removed.edges[seq(length=i)])
  cl <- clusters(g2)$membership
# March 13, 2014 - compute modularity on the original graph g 
# (Thank you to Augustin Luna for detecting this typo) and not on the induced one g2. 
  modularity(g9,cl)
})
 
# we can now plot all modularities
plot(mods, pch=20)

 
```

```{r}
ebc$removed.edges
```


```{r, fig.width = 12, fig.height = 12}
# Now, let's color the nodes according to their membership
#g2<-delete.edges(g9, ebc$removed.edges[seq(length=which.max(mods)-1)])
g2<-delete.edges(g9, ebc$removed.edges[seq(length=4)])
V(g9)$color <- clusters(g2)$membership
 
table(V(g9)$color)
as.factor(clusters(g2)$membership)

# plot it
plot(g9, vertex.label=NA, vertex.label=NA, vertex.size=2, vertex.col=as.factor(clusters(g2)$membership))
```

